Let us consider the following optimization problem:
\begin{subequations}\label{eq:problem} 
\begin{align}
\min_{x,y} & \quad f(x,y)  \\
\text{s.t.} & \quad l \leq x \leq u \\
& \quad \mathds{1}^T x = 1 
\end{align}
\end{subequations}
where $x \in \R^n$, $y \in \R^m$, $f$ continuously differentiable, $l, u \in \R^n$ with $l < u$ and $\mathds{1} \in \R^n$ is all composed by ones. 

We define the feasible set $\mathcal{F}$  of Problem (\ref{eq:problem}):
\begin{equation}
\mathcal{F} = \{(x,y) \in \R^{n+m} : \mathds{1}^T x = 1, l \leq x \leq u\}.
\end{equation}
A vector $d\in \R^{n+m}$ is partitioned as follows
$$
d=\left(
\begin{array}{c}
d_x\\
d_y
\end{array}
\right ),
$$
where $d_x\in R^n$ and $d_y\in R^m$.

Given $(x,y) \in \mathcal{F}$, the set of feasible directions in $(x,y)$ is the cone
\begin{equation}
 \mathcal{D}(x,y)=\{ d \in \R^{n+m}: \mathds{1}^Td_x=0, d_i\ge 0 \ \forall i \in L(x), d_i\le 0 \ \forall i \in U(x)\}
\end{equation}
where
\begin{equation}
 \begin{aligned}
  &L(x)=\{ i: \ x_i=l_i\}\\
  &U(x)=\{ i: \ x_i=u_i\}
 \end{aligned}
\end{equation}
Given $(\bar x,\bar y) \in \mathcal{F}$, we say that $(\bar x,\bar y)$ is a {\it critical point} if
$$
\nabla f(\bar x,\bar y)^Td\ge 0\quad\quad \forall d\in  \mathcal{D}(\bar x,\bar y).
$$
We can state the following result.

\begin{proposition}[Name]
\label{optimality}
A point $(\bar x,\bar y) \in \mathcal{F}$ is a critical point if and only if
\begin{equation}\label{on_x}
\begin{aligned}
&\nabla_xf(\bar x,\bar y)^Td_x\ge 0 \quad \forall d_x\in R^n & \\ 
&\text{s.t.} \quad \mathds{1}^Td_x=0,\quad d_i\ge 0 \ \forall i \in L(\bar x), \quad d_i\le 0 \ \forall i \in U(\bar x)&
\end{aligned}
\end{equation}
\begin{equation}\label{on_y}
 \nabla_yf(\bar x,\bar y)=0.
\end{equation} 
\end{proposition}


In correspondence to a feasible point $(x,y)$ we introduce the index sets
$$
R(x)=L(x)\cup C(x)
$$
$$
S(x)=U(x)\cup C(x),
$$
where 
$$
C(x)=\{i: l_i<x_i<u_i\}.
$$
We can state the following propositions (Propositions 2.2 and 2.3 in Jota2009).
\begin{proposition}\label{2.2}
 A feasible point $(\bar x,\bar y)$ is a critical point if and only if for each pair $(i,j)$,
$i\in R(\bar x)$, $j\in S(\bar x)$, we have
\begin{equation}\label{on_RS}
 {{\partial f(\bar x,\bar y)}\over{\partial x_i}}\ge
 {{\partial f(\bar x,\bar y)}\over{\partial x_j}}
\end{equation}
\begin{equation}\label{on_y2}
\nabla_yf(\bar x,\bar y)=0.
\end{equation}
\end{proposition}
\begin{proposition}\label{2.3}
 Let $\{(x^k,y^k)\}$ be a sequence of feasible points convergent to a point $(\bar x,\bar y)$.
Then, for sufficiently large values of $k$, we have
$$
R(\bar x)\subseteq R(x^k) \quad \quad {\rm and}\quad \quad S(\bar x)\subseteq S(x^k).
$$
\end{proposition}

\subsection{Set of sparse feasible directions}
In our decomposition framework we will employ feasible directions having only two
nonzero components.
Then, in this subsection we introduce these sparse feasible directions and we show their important properties.

Given $i, j\in  \{1, \ldots ,n\}$, with $i\ne j$,
we indicate by $d^{i,j}\in  \R^{n+m}$ such that
\begin{equation}\label{eq:direction}
d_h^{i,j}= 
\begin{cases}
1, \quad \text{    } h=i\\
-1, \text{    } \text{    } h=j\\
0, \quad \text{    } \text{otherwise}
\end{cases}
\end{equation}
Note that, by definition, the components $d_h^{i,j}$ with $h=n+1,\ldots,n+m$ are always set to zero.

Given $(x, y) \in \mathcal{F}$ and the corresponding index sets $R(x)$ and $S(x)$, we indicate by $D_{RS}(x,y)$
the set of directions $d^{i,j}$ with $i \in R(x)$ and $j \in S(x)$, namely
$$
D_{RS}(x,y)=\cup_{i\in R(x),j\in S(x)}d^{i,j}.
$$
\begin{proposition}\label{3.1}
Let $(\bar x,\bar y)$ be a feasible point. For each pair $i \in R(x)$ and $j \in S(x)$, the
direction $d^{i,j}\in \R^{n+m}$ is a feasible direction at $(\bar x,\bar y)$, i.e. $d \in D(\bar x,\bar y)$.
\end{proposition}
\begin{proposition}\label{3.2}
A feasible point $(\bar x,\bar y)$
 is a critical point if and only
\begin{equation}\label{on_x2}
\nabla f(\bar x,\bar y)^Td^{i,j}\ge 0\quad\quad \forall d^{i,j}\in D_{RS}(\bar x,\bar y)
\end{equation}
\begin{equation}\label{on_y3}
 \nabla_y f(\bar x,\bar y)=0.
\end{equation} 
\end{proposition}
Given a feasible point $(\bar x,\bar y)$, a pair $i\in R(\bar x)$ and $j\in S(\bar x)$ such that
$$
\nabla f(\bar x,\bar y)^Td^{i,j}<0
$$
is said a {\it Violating Pair} (VP).

A violating pair $(i^\star,j^\star)$ such that 
\begin{equation}\label{mvp}
 \nabla f(\bar x,\bar y)^Td^{i^\star,j^\star}\le \nabla f(\bar x,\bar y)^Td^{i,j} \quad \forall i\in R(\bar x), \ j\in S(\bar x).
\end{equation}
is the {\it Most Violating Pair} (MVP).


